---
title: "STAT 312R - HW1"
author: Vinayak Mathur
date: 27th January 2022
output:
pdf_document: default
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1:
An article in a women's magazine reported that women who nurse their babies feel warmer and
more receptive towards their infants than mothers who bottle-feed. This conclusion was based
on the following scores from a questionnaire designed to measure warmth and receptivity,
given to a random sample of 10 women from each group taken from recent hospital birth
records (higher score represents more warmth and reception).\
Nurse:  48.4, 51.5, 45.0, 47.0, 46.9, 47.7, 49.9, 48.5, 45.6, 48.9\
Bottle feed: 42.5, 43.8, 44.5, 47.5, 44.5, 45.6, 43.4, 44.6, 47.1, 45.8\
a) Is this an observational study or a controlled experiment?  Justify your answer. (2 pt)\
b) Using SOCS, summarize the data for the two groups. (4 pts)\
c) Using your descriptions in part b), compare the scores in the two groups. (2 pts)\
d) If you or your partner were thinking about breast-feeding, what effect would this study have
on your decision? Why? (2 pts)\
```{r}
nurse <- c(48.4, 51.5, 45.0, 47.0, 46.9, 47.7, 49.9, 48.5, 45.6, 48.9)
bottle <- c(42.5, 43.8, 44.5, 47.5, 44.5, 45.6, 43.4, 44.6, 47.1, 45.8)
summary(nurse)
summary(bottle)
boxplot(nurse)
boxplot(bottle)
hist(nurse,
     main="Histogram for Nursed Infants",
     xlab="Warmth and Reception",
     border="green",
     col="blue")
hist(bottle,
     main="Histogram for Bottle-fed Infants",
     xlab="Warmth and Reception",
     border="green",
     col="pink")

```
ans) From the above calculations:\
a) This is an observational study.\
b) Nursed: Shape -> uni-modal distribution skewed to the left\
Outliers -> No outliers\
Center , Spread-> \
Min. 1st Qu.  Median    Mean 3rd Qu.    Max.\
45.00   46.92   48.05   47.94   48.80   51.50\
\
Nursed: Shape -> uni-modal distribution skewed to the right\
Outliers -> No outliers\
Center , Spread-> \
Min. 1st Qu.  Median    Mean 3rd Qu.    Max.\
42.50   43.98   44.55   44.93   45.75   47.50\

## Problem 2:\
One of the authors (DCM) has a Mercedes-Benz 500 SL Roadster.
It is a 2003 model and has fairly low mileage (currently
60,112 miles on the odometer). He is interested in learning how
his car’s mileage compares with the mileage on similar SLs.
The table below contains the mileage on 100 Mercedes-Benz
SLs from the model years 2003–2009 taken from the Cars.com
website.\
### a)\
Calculate the sample mean and standard deviation of the odometer readings.\
### b)\
Construct a histogram of the odometer readings and comment on the shape of the data distribution.\
### c)\
Construct a stem-and-leaf diagram of the odometer readings.\
### d)\
What is the percentile of DCM’s mileage?\
```{r}
miles <- read.table("MercMileage.txt")$V1
dcm <- 60112
mean(miles)
sd(miles)
hist(miles,
     main="Histogram for Mercedes Mileage",
     xlab="Miles",
     border="green",
     col="blue")
stem(miles, scale = 2)
plot(ecdf(miles))
ecdf(miles)(dcm)

```
ans) From the above calculations:\
a) mean = 34232.05 miles\
standard deviation = 20414.52\
d) 84%