---
title: "Discrete Probability Distributions"
author: "Anirban Mondal"
date: "01/13/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Binomial Distribution

Find probabilities related to this distribution
Example: Let X be binomial random variable with n=10 and p=0.5.
```{r}
?dbinom ## pmf for binomial distribution
n=10
p=0.5
dbinom(2,n,p) ## Finds P(X=2) 
pbinom(2,n,p) ## Finds P(X <= 2) = P(X=0)+P(X=1)+P(X=2)
mu = n*p ## mean
s2 = n*p*(1-p) ## variance
mu
s2
```

Plotting the pmf and cdf  of Binomial distribution
```{r}
n=5
p=0.5
xx=0:n
## Plotting pmf 
plot(xx,dbinom(xx,n,p),type="h",col="blue",lwd=2,main="PMF for Binomial 
     distribution with n=5, p=0.5",xlab="x",ylab="f(x)")
points(xx,dbinom(xx,n,p), pch=20, cex=2, col="blue")

## Plotling CDF 
yy=c(0,pbinom(xx,n,p))
plot(stepfun(xx,yy),vertical=F, pch=20, cex=2, lwd=2, col="blue", ylab="F(x)",
     main ="CDF for Binomial distribution with n=5, p=0.5")
grid()
```

Simulation from binomial distribution

```{r}
rbinom(10,n,p) # Geneartes 10 random samples from a binomial distribution
ns=1000 ## sample size
set.seed(20) ## setting seed fixes the random number generator
x= rbinom(ns,n,p) 
tx=table(x) ## Frequency table for the sample
barplot(table(x)) ## barplot for the frequency
barplot(prop.table(table(x))) ## barplot using probabilities

### Plotting emperical cdf of the simulated data and the theoretical cdf overlay
## ecdf (F_n(x)): (Number of x_i <=x)/ns
plot(ecdf(x), lwd=2, pch=20, cex=2, main ="CDF for Binomial distribution 
     with n=5, p=0.5", ylab="F(x)")
yy=c(0,pbinom(0:n,n,p))
lines(stepfun(0:n,yy),vertical=F, pch=20, cex=2, lwd=2, col="blue")
legend("bottomright", c("Emperical cdf", "Theoretical cdf"), lty=c(1,1), 
       lwd=c(2,2), col=c("black", "blue"))
```



Quantile function for binomial distribution
Quantile Definition: $Q(r)=inf\{x \in R: F(x) \geq r\}$
```{r}
qbinom(0.1, n,p) ## Finds the value smallest x such that P(X<=x)>=0.1 (quantile)
pbinom(2,n,p)
pbinom(3,n,p)
```

## Geometric distribution
Note that the geometric distribution in R is defined as the number of failures,
not the number of trails. 
X: no of failures before getting the 1st success. 
X=Y-1, where Y is the number of trails to get the first success. 
E(X)=E(Y)-1=1-1/p, V(Y)=V(X)
Example: X follows a gemetric distribution of success probablity p=0.2
```{r}
?dgeom
p=0.2
dgeom(3,p) ## P(X=3)
pgeom(2,p) ## P(X<=2)
rgeom(10,p) ## Generate 10 random samples from geometric distribution
qgeom(0.25, p) ## 25th quantile or 1st quartile of the geometric distribution
## min x such that P(X<=x)>=0.25
```

Plotting the pmf and cdf  of Geometric distribution
```{r}
p=0.2
xend=10
xx=0:xend 
## Plot pmf 
plot(xx,dgeom(xx,p),type="h",col="blue",lwd=2,main="PMF for Geomteric 
     distribution with p=0.2",xlab="x",ylab="f(x)")
points(xx,dgeom(xx,p), pch=20, cex=2, col="blue")
## Plot CDF 
yy=c(0,pgeom(xx,p))
plot(stepfun(xx,yy),vertical=F, pch=20, cex=2, lwd=2, col="blue", 
     ylab="F(x)", main ="CDF for Geometric distribution with p=0.2", xlim=c(xx[1], xend))
grid()
```


Simulation from Geometric distribution

```{r}
p=0.2
rgeom(10,p) # Geneartes 10 random samples from a Geometric distribution
ns=100 ## sample size
set.seed(20) ## setting seed fixes the random number generator
x2= rgeom(ns,p) 
table(x2) ## Frequency table for the sample
barplot(table(x2)) ## barplot for the frequency
barplot(prop.table(table(x2)) ) ## barplot using percentages

### Plotting emperical cdf of the simulated data and the theoretical cdf overlay
xx=0:max(x2) 
yy=yy=c(0,pgeom(xx,p))
plot(ecdf(x2), lwd=2, pch=20, cex=2, main ="CDF for Geometric distribution
     with p=0.2", ylab="F(x)", xlim=c(xx[1], max(x2)))
lines(stepfun(xx,yy),vertical=F, pch=20, cex=2, lwd=2, col="blue")
legend("bottomright", c("Emperical cdf", "Theoretical cdf"), lty=c(1,1), 
       lwd=c(2,2), col=c("black", "blue"))
```


## Hypergeometric Distribution

Find probabilities related to this distribution
Example: Let X be Hypergeometric random variable with N=20, K=8, n=5.
N: Total number of balls (m+n in R)
K: Number of white balls (m in R)
n: Number of balls drawn without replacment (k in R)
```{r}
?dhyper ## pmf for Hypergeometric distribution
n=5
N=20
K=8
dhyper(2,K,N-K,n) ## Finds P(X=2)  dhyper(x,m,n,k)
phyper(2,K,N-K,n) ## Finds P(X <= 2) = P(X=0)+P(X=1)+P(X=2)
mu = n*K/N ## mean
s2 = n*K/N*(N-K)/N*(N-n)/(N-1) ## variance
mu
s2
```

Plotting the pmf and cdf  of Hyergeometric distribution
```{r}
xx=0:n
## Plotting pmf 
plot(xx,dhyper(xx,K,N-K,n),type="h",col="blue",lwd=2,main="PMF for Hypergeometric
     distribution",xlab="x",ylab="f(x)")
points(xx,dhyper(xx,K,N-K,n), pch=20, cex=2, col="blue")

## Plotling CDF 
yy=c(0,phyper(xx,K,N-K,n))
plot(stepfun(xx,yy),vertical=F, pch=20, cex=2, lwd=2, col="blue", ylab="F(x)",
     main ="CDF for Hypergeometric distribution")
grid()
```

Simulation from Hypergeometric distribution

```{r}
rhyper(10,K,N-K,n) # Geneartes 10 random samples from a Hypergeometric distribution
ns=100 ## sample size
set.seed(20) ## setting seed fixes the random number generator
x= rhyper(ns,K,N-K,n) 
tx=table(x) ## Frequency table for the sample
barplot(table(x)) ## barplot for the frequency
barplot(prop.table(table(x))) ## barplot using probabilities

### Plotting emperical cdf of the simulated data and the theoretical cdf overlay
## ecdf (F_n(x)): (Number of x_i <=x)/ns
plot(ecdf(x), lwd=2, pch=20, cex=2, main ="CDF for Hypergeometric distribution",
     ylab="F(x)")
xx=0:n
yy=c(0,phyper(0:n,K,N-K,n))
lines(stepfun(0:n,yy),vertical=F, pch=20, cex=2, lwd=2, col="blue")
legend("bottomright", c("Emperical cdf", "Theoretical cdf"), lty=c(1,1), 
       lwd=c(2,2), col=c("black", "blue"))
```



Quantile function for Hypergeometric distribution
Quantile Definition: $Q(r)=inf\{x \in R: F(x) \geq r\}$
```{r}
qhyper(0.5, K,N-K,n) ## Finds the value smallest x such that P(X<=x)>=0.1 (quantile)
phyper(1,K,N-K,n)
phyper(2,K,N-K,n)
```

## Poisson distribution

Example: X follows a Poisson distribution of lambda=2 (mean)
```{r}
?dpois
l=2
dpois(3,l) ## P(X=3)
ppois(2,l) ## P(X<=2)
rpois(10,l) ## Generate 10 random samples from Poisson distribution
qpois(0.25, l) ## 25th quantile or 1st quartile of the Poisson distribution
## min x such that P(X<=x)>=0.25
```

Plotting the pmf and cdf  of Poisson distribution
```{r}
l=2
xend=10
xx=0:xend 
## Plot pmf 
plot(xx,dpois(xx,l),type="h",col="blue",lwd=2,main="PMF for Poisson
     distribution",xlab="x",ylab="f(x)")
points(xx,dpois(xx,l), pch=20, cex=2, col="blue")
## Plot CDF 
yy=c(0,ppois(xx,l))
plot(stepfun(xx,yy),vertical=F, pch=20, cex=2, lwd=2, col="blue", 
     ylab="F(x)", main ="CDF for Poisson distribution", 
     xlim=c(xx[1], xend))
grid()
```


Simulation from Poisson distribution

```{r}
l=2
rpois(10,l) # Geneartes 10 random samples from a Poisson distribution
ns=1000 ## sample size
set.seed(20) ## setting seed fixes the random number generator
x2= rpois(ns,l) 
table(x2) ## Frequency table for the sample
barplot(table(x2)) ## barplot for the frequency
barplot(prop.table(table(x2)) ) ## barplot using percentages

### Plotting emperical cdf of the simulated data and the theoretical cdf overlay
xx=0:max(x2) 
yy=yy=c(0,ppois(xx,l))
plot(ecdf(x2), lwd=2, pch=20, cex=2, main ="CDF for Poisson distribution",
     ylab="F(x)", xlim=c(xx[1], max(x2)))
lines(stepfun(xx,yy),vertical=F, pch=20, cex=2, lwd=2, col="blue")
legend("bottomright", c("Emperical cdf", "Theoretical cdf"), lty=c(1,1), 
       lwd=c(2,2), col=c("black", "blue"))
```
