---
title: "STAT 312R - HW2"
author: Vinayak Mathur
date: 23rd February 2022
output:
pdf_document: default
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Information on a packet of seeds claims that 42% of seed in the package produce large tomatoes, while 38% produce yellow tomatoes.  Assume that size and color are independent. What’s the probability that a given seed will produce a large tomato or a yellow tomato?\
ans) Let L be the event of getting a large tomato\
Let Y be the event of a yellow tomato\
P(L) = 0.42\
P(Y) = 0.38\
=> P(L U Y) = P(L) + P(Y) - P(L n Y)\
=> P(L U Y) = 0.42 + 0.38 - 0.42*0.38\
```{r}
print(0.42 + 0.38 - 0.42 * 0.38)
```

2. In a NiCd battery, a fully charged cell is composed of nickelic hydroxide. Nickel is an element that has multiple oxidation states. Assume the following proportions of the states:\
Nickel Charge      Proportions Found\
0       0.17\
+2      0.35\
+3      0.33\
+4      0.15\
a) Determine the cumulative distribution function of the nickel charge.\
```{r}
nickelCharge <- c(0, 2, 3, 4)
proportionsFound <- c(0, 0.17, 0.35, 0.33, 0.15)

## Plotling CDF
plot(stepfun(nickelCharge, cumsum(proportionsFound)), vertical = F, pch = 20, cex = 2, lwd = 2, col = "blue", ylab = "F(x)",
     main = "CDF for the nickel charge")
grid()
```

b) Determine the mean and variance of the nickel charge.\
```{r}
mean(nickelCharge)
sd(nickelCharge)^2
```
3. Heart failure is due to either natural occurrences (87%) or outside factors (13%). Outside factors are related to induced substances or foreign objects. Natural occurrences are caused by arterial blockage, disease, and infection. Suppose
that 20 patients will visit an emergency room with heart failure. Assume that causes of heart failure for the individuals are independent.\
a. What is the probability that three individuals have conditions caused by outside factors?\
```{r}
p <- 0.13
ans1 <- dbinom(3, 20, p)
print(ans1)
```
b. What is the probability that three or more individuals have conditions caused by outside factors?\
```{r}
print(1 - dbinom(2, 20, p))
```
c. What are the mean and standard deviation of the number of individuals with conditions caused by outside factors?\
```{r}
res <- 20 * p
print(res)
print(sqrt(20 * 0.13 * 0.87))
```

6. The demand for water use in Phoenix in
2003 hit a high of about 442 million gallons per day on June 27
(http://phoenix.gov/WATER/wtrfacts.html). Water use in the sum-
mer is normally distributed with a mean of 310 million gallons per
day and a standard deviation of 45 million gallons per day. City
reservoirs have a combined storage capacity of nearly 350 million
gallons.
a. What is the probability that a day requires more water
than is stored in city reservoirs?

```{r}
qnorm(0.75,5000, 1000)

```
b. What reservoir capacity is needed so that the probability
that it is exceeded is 1%?
```{r}
qnorm(0.99, 310, 45)
qnorm(0.99, 11, 0.3)

```
c. What amount of water use is exceeded with 95%
probability?
```{r}
qnorm(0.99, 310, 45)

```
d. Water is provided to approximately 1.4 million people.
What is the mean daily consumption per person at which
the probability that the demand exceeds the current reser-
voir capacity is 1%? Assume that the standard deviation of
demand remains the same