---
title: "Quiz 1"
author: "Vinayak Mathur"
date: "2/1/22"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1: 
Disks of polycarbonate plastic from a supplier are analyzed for scratch and shock resistance. The results from 100 disks are summarized as follows:

            Shock Resistance\
                      High(A)     Low\
Scratch     High(B)   70          9\
Resistance  Low       16          5\

Let A denote the event that a disk has high shock resistance, and let B denote the event that a disk has high scratch resistance. If a disk is selected at random, determine the following probabilities: (6 points)

### a) 
P(A)

```{r}

```
Ans: P(A) = 0.86

### b) 
P(B)

```{r}

```
Ans: P(B) = 0.79

### c)
P(Ac)

```{r}

```
Ans: P(Ac)  = 0.14

### d)
P(A n B)

```{r}

```
Ans: P(A n B)  = 0.7

### e)
P(A U B)

```{r}

```
Ans: P(A U B)   = 0.95

### f)
P(Ac U B)

```{r}

```
Ans: P(Ac U B)   = 0.84