---
title: "R Lab 4"
author: "Vinayak Mathur"
date: "5th April 2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1:

A random sample of 16 Americans in a city yielded the following data on the number of pounds of beef consumed per year: 118 ,115.125,110, 112, 130,117 ,112 ,115 ,120, 113, 118, 119, 122, 123, 126.


### a) (4 points)
Construct a quantile-quantile plot for the data to check if the annual beef consumption in the given city can be modelled by a normal distribution.

```{r}
sampleData <- c(118, 115, 125, 110, 112, 130, 117, 112, 115, 120, 113, 118, 119, 122, 123, 126)
qqnorm(sampleData, pch = 1, frame = FALSE)
qqline(sampleData, col = "steelblue", lwd = 2)
shapiro.test(sampleData)
```
Ans:

### b) (4+2=6 points)
Construct a two-sided 95% confidence interval for the mean beef consumption of the people in the given city and interpret it.\

95% two-sided CI where neither $\mu$ nor $\sigma$ are known is given by $(\bar{X} - t_{\alpha/2, n-1}\displaystyle\frac{s}{\sqrt(n)}, \bar{X} + t_{\alpha/2, n-1}\frac{s}{\sqrt(n)})$\
```{r}

t.test(sampleData)
x_bar <- mean(sampleData)
s <- sd(sampleData)
n <- length(sampleData)
alpha <- 0.95
t <- qt((1-alpha)/2, n-1)
LCL <- x_bar - abs(t*s/sqrt(n))
UCL <- x_bar + abs(t*s/sqrt(n))
c(LCL, UCL)



```

