---
title: "R Lab 2"
author: "Vinayak Mathur"
date: "1/25/22"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1: 
The energy consumption for 90 gas-heated homes during a winter heating season is given in the file "Energy_Usage.csv" (see Data module on canvas). The variable reported is BTU/number of heating degree day. 

### a) 
Use R codes to find the five point summary, mean and standard deviation of the data. (4 points)

```{r}
energy <- read.csv("Energy_Usage.csv", header = TRUE)
summary(energy)
sd(energy$BTU)
```
Ans: Standard Deviation = 2.86799

### b) 
Construct a histogram and a stem-and-leaf diagram of energy usage. Use appropriate scale for the stem-and-leaf diagram. Then use SOCS method to describe the data. (6 points)

```{r}
hist(energy$BTU,
     main="Histogram for Energy usage of 90 homes",
     xlab="Energy Usage in BTU",
     border="green",
     col="blue")
stem(energy$BTU, scale = 2)
summary(energy$BTU)
```
Ans: Shape -> Shape of the distribution is symmetric with a single peak.
Outliers -> The data does not have any outliers.
Min. 1st Qu.  Median    Mean 3rd Qu.    Max.
2.970   7.947   9.835  10.038  12.045  18.260


