---
title: "STAT 312R - HW5"
author: Vinayak Mathur
date: 20st April 2022
output:
pdf_document: default
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Question 2 8.S10

During the 1999 and 2000 baseball seasons, there was
much speculation that the unusually large number of home runs
hit was due at least in part to a livelier ball. One way to test
the “liveliness” of a baseball is to launch the ball at a vertical
surface with a known velocity VL and measure the ratio of the
outgoing velocity VO of the ball to VL. The ratio $R = VO/VL$ is
called the coefficient of restitution. Following are measurements
of the coefficient of restitution for 40 randomly selected base-
balls. The balls were thrown from a pitching machine at an oak
surface. \
\

(0.6248, 0.6237, 0.6118, 0.6159, 0.6298, 0.6192, 0.6520, 0.6368, 0.6220, 0.6151, 0.6121, 0.6548, 0.6226, 0.6280, 0.6096, 0.6300, 0.6107, 0.6392, 0.6230, 0.6131, 0.6223, 0.6297, 0.6435, 0.5978, 0.6351, 0.6275, 0.6261, 0.6262, 0.6262, 0.6314, 0.6128, 0.6403, 0.6521, 0.6049, 0.6170, 0.6134, 0.6310, 0.6065, 0.6214, 0.6141)
\
\
a. Is there evidence to support the assumption that the\
coefficient of restitution is normally distributed?\

```{r}
coeffs <- c(0.6248, 0.6237, 0.6118, 0.6159, 0.6298, 0.6192, 0.6520, 0.6368, 0.6220, 0.6151, 0.6121, 0.6548, 0.6226, 0.6280, 0.6096, 0.6300, 0.6107, 0.6392, 0.6230, 0.6131, 0.6223, 0.6297, 0.6435, 0.5978, 0.6351, 0.6275, 0.6261, 0.6262, 0.6262, 0.6314, 0.6128, 0.6403, 0.6521, 0.6049, 0.6170, 0.6134, 0.6310, 0.6065, 0.6214, 0.6141)

library(qualityTools)
qqPlot(coeffs, "Normal")
```


b. Find a 99% CI on the mean coefficient of restitution.\
```{r}

n <- length(coeffs)
xbar <- mean(coeffs)
s <- sd(coeffs)
t <- qt(p = .005, df = n - 1)
margin <- t * s / sqrt(n)
lower <- xbar + margin
upper <- xbar - margin

print(lower)
print(upper)

```
Ans) $0.6187462 \leq \mu \leq 0.6299288$\



c. Find a 99% prediction interval on the coefficient of\
restitution for the next baseball that will be tested.\
```{r}
margin <- t * s * sqrt(n + 1) / sqrt(n)
lower <- xbar + margin
upper <- xbar - margin

print(lower)
print(upper)

```
Ans) $0.0.5885355 \leq \mu \leq 0.6601395$\


### Question 4 9.2.8

Supercavitation is a propulsion technology for undersea vehicles that can greatly increase their speed. It occurs above
approximately 50 meters per second when pressure drops sufficiently to allow the water to dissociate into water vapor, forming
a gas bubble behind the vehicle. When the gas bubble completely encloses the vehicle, supercavitation is said to occur.
Eight tests were conducted on a scale model of an undersea
vehicle in a towing basin with the average observed speed
$x = 102.2$ meters per second. Assume that speed is normally
distributed with known standard deviation $\sigma = 4$ meters per
second.\


a. Test the hypothesis $H0: \mu = 100$ versus $H1: \mu < 100$ using
$α = 0.05$.
```{r}
mu <- 100
xbar <- 102.2
sig <- 4
n <- 8

z0 <- (xbar - mu) / (sig / sqrt(n))

za <- 1.65
# we reject H0 if z0<-za
res <- z0 < (-za)
print(res)



```
Ans) We do not reject $H_0$ and reject $H_1$\





b. What is the P-value for the test in part (a)?
```{r}
2 * pnorm(-abs(z0))
```

c. Compute the power of the test if the true mean speed is
as low as 95 meters per second.


```{r}
mu2 <- 95
z0 <- (mu2 - mu) / (sig / sqrt(n))
print(z0)

print(1 - pnorm(-za - z0))

```
ans) 0.0296789\


### Question 5 9.3.11


Given ASU female student height data in inches: (62, 64, 61, 67, 65, 68, 61, 65, 60, 65, 64, 63, 59, 68, 64, 66, 68, 69, 65, 67, 62, 66, 68, 67, 66, 65, 69, 65, 69, 65, 67, 67, 65, 63, 64, 67, 65)\

$H_0: \mu \geq 65$\
$H_1: \mu < 65$\

a. Can you support a claim that the mean height of female
engineering students at ASU is at least 65 inches? Use
$\alpha = 0.05$. Find the P-value.

```{r}
heightData <- c(62, 64, 61, 67, 65, 68, 61, 65, 60, 65, 64, 63, 59, 68, 64, 66, 68, 69, 65, 67, 62, 66, 68, 67, 66, 65, 69, 65, 69, 65, 67, 67, 65, 63, 64, 67, 65)


mu <- 65

t.test(heightData, mu = mu)
qt(0.05/2, 36, lower.tail = FALSE)
```
ans) Since 0.3877 < 2.028, reject the alternate hypothesis and conclude that there is sufficient evidence to indicate that the true mean height of female engineering students is equal to 65 at α = 0.05\




b. Check the normality assumption.
```{r}
qqPlot(heightData, "Normal")
shapiro.test(heightData)
```
ans) Since p-value = 0.1053 > 0.05, $H_0$ is accepted and the distribution is concluded to be normally distributed.\




c. Compute the power of the test if the true mean height is
68 inches.
```{r}
t.test(heightData, mu = 68)
1-6.299e-08
```
ans) p-value = 1\


